Reasoning About a Capability Machine with Local Capabilities

### 2017-09-28 esop-2018/paper.tex
removed part of the example section, namely the world construction for the last proof.

In Lemma~\ref{lem:correctness-g1} the order of control has been
inverted compared to the previous lemmas. In this lemma, the adversary
assumes control first with a capability for the closure creator
\texttt{\footnotesize{g1}}. Consequently, we need to check that all
arguments are safe to use and that we clean up before returning in the
end. The inversion of control poses an interesting challenge when it
comes to reasoning about the adversary's local state during the
execution of \texttt{\footnotesize{f4}} and the callbacks where the
adversary should not rely on the local state from before the call of
\texttt{\footnotesize{f4}}. This is easily done by revoking all the
temporary regions of the world given at the start of
\texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
returns, the adversary is again allowed to rely on its old local state
so we need to guarantee that the local state is unchanged. This is
important because the return pointer that \texttt{\footnotesize{f4}}
receives may be local, and the adversary is allowed
to allocate the activation record on the stack (just like we do) so
they can store and recover their old stack pointer after
\texttt{\footnotesize{f4}} returns.

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
for all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{\footnotesize{f4}} from the adversary
is safe in $W_1$, so it can only be used in public future worlds. Let us see how
$W_6$ is constructed as a public future world of $W_1$, as well as a private
future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
away and static temporary regions (a region that accepts exactly one memory) for
our local stack as well as the adversary's local stack. The given worlds $W_3$
and $W_5$ are public future worlds in which the adversary chooses to return to
us. None of these worlds can have masked the temporary regions of $W_1$ with
permanent regions. $W_6$ is constructed by revoking the temporary regions of
$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
future world of $W_1$. It is therefore safe to use the return pointer to return
to the adversary, since we have restored validity of any local state they might
have stashed away.

@@

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f4} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
\caption{Illustration of the worlds in the proof of
  Lemma~\ref{lem:correctness-g1}. In the proof, the top row of worlds are
  constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{comment}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?

>>>>

In Lemma~\ref{lem:correctness-g1} the order of control has been
inverted compared to the previous lemmas. In this lemma, the adversary
assumes control first with a capability for the closure creator
\texttt{\footnotesize{g1}}. Consequently, we need to check that all
arguments are safe to use and that we clean up before returning in the
end. The inversion of control poses an interesting challenge when it
comes to reasoning about the adversary's local state during the
execution of \texttt{\footnotesize{f4}} and the callbacks where the
adversary should not rely on the local state from before the call of
\texttt{\footnotesize{f4}}. This is easily done by revoking all the
temporary regions of the world given at the start of
\texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
returns, the adversary is again allowed to rely on its old local state
so we need to guarantee that the local state is unchanged. This is
important because the return pointer that \texttt{\footnotesize{f4}}
receives may be local, and the adversary is allowed
to allocate the activation record on the stack (just like we do) so
they can store and recover their old stack pointer after
\texttt{\footnotesize{f4}} returns. By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world which the adversary's invariants are preserved. The details of this and the proofs of the other lemmas are found in the technical appendix\cite{technical_appendix}.

\begin{comment}
In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
for all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{\footnotesize{f4}} from the adversary
is safe in $W_1$, so it can only be used in public future worlds. Let us see how
$W_6$ is constructed as a public future world of $W_1$, as well as a private
future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
away and static temporary regions (a region that accepts exactly one memory) for
our local stack as well as the adversary's local stack. The given worlds $W_3$
and $W_5$ are public future worlds in which the adversary chooses to return to
us. None of these worlds can have masked the temporary regions of $W_1$ with
permanent regions. $W_6$ is constructed by revoking the temporary regions of
$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
future world of $W_1$. It is therefore safe to use the return pointer to return
to the adversary, since we have restored validity of any local state they might
have stashed away.

@@

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f4} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
\caption{Illustration of the worlds in the proof of
  Lemma~\ref{lem:correctness-g1}. In the proof, the top row of worlds are
  constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\end{comment}
\begin{comment}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?
